---
title: "Assignment_2"
author: "Saurav Singh"
date: "28 February 2017"
output: pdf_document
---

```{r}
par(mfrow=c(2,2))
curve(1 - log(x) * exp(-x^2),0.8,2)  
```  


**1. Unconstrained optimization in one dimension (f(x) =1 - log(x) * exp(-x^2).)**  

**Newton's Method**  
```{r}
newton <- function(f, tol=1E-12,x0=1,N=20) {
  h <- 0.001
  i <- 1; x1 <- x0
  p <- numeric(N)
  while (i<=N) {
    df.dx <- (f(x0+h)-f(x0))/h
    x1 <- (x0 - (f(x0)/df.dx))
    p[i] <- x1
    i <- i + 1
    if (abs(x1-x0) < tol) break
    x0 <- x1
  }
  return(p[1:(i-1)])
}
f <-function(x) { {4*x^2 + 5 * x^3} }
p <- newton(f, x0=1, N=10)
p
```
**Golden-Section Search**
```{r}
f <- function(x) {1 - log(x) * exp(-x^2)}
golden.section.search = function(f, a, b, tolerance)
{
  golden.ratio = 2/(sqrt(5) + 1)
  
  ### Use the golden ratio to set the initial test points
  x1 = b - golden.ratio*(b - a)
  x2 = a + golden.ratio*(b - a)
  
  ### Evaluate the function at the test points
  f1 = f(x1)
  f2 = f(x2)
  
  iteration = 0
  
  while (abs(b - a) > tolerance)
  {
    iteration = iteration + 1
    cat('', '\n')
    cat('Iteration #', iteration, '\n')
    cat('f1 =', f1, '\n')
    cat('f2 =', f2, '\n')
    
    if (f2 > f1)
      # then the minimum is to the left of x2
      # let x2 be the new upper bound
      # let x1 be the new upper test point
    {
      cat('f2 > f1', '\n')
      ### Set the new upper bound
      b = x2
      cat('New Upper Bound =', b, '\n')
      cat('New Lower Bound =', a, '\n')
      ### Set the new upper test point
      ### Use the special result of the golden ratio
      x2 = x1
      cat('New Upper Test Point = ', x2, '\n')
      f2 = f1
      
      ### Set the new lower test point
      x1 = b - golden.ratio*(b - a)
      cat('New Lower Test Point = ', x1, '\n')
      f1 = f(x1)
    } 
    else 
    {
      cat('f2 < f1', '\n')
      # the minimum is to the right of x1
      # let x1 be the new lower bound
      # let x2 be the new lower test point
      
      ### Set the new lower bound
      a = x1
      cat('New Upper Bound =', b, '\n')
      cat('New Lower Bound =', a, '\n')
      
      ### Set the new lower test point
      x1 = x2
      cat('New Lower Test Point = ', x1, '\n')
      
      f1 = f2
      
      ### Set the new upper test point
      x2 = a + golden.ratio*(b - a)
      cat('New Upper Test Point = ', x2, '\n')
      f2 = f(x2)
    }
  }
  
  ### Use the mid-point of the final interval as the estimate of the optimzer
  cat('', '\n')
  cat('Final Lower Bound =', a, '\n')
  cat('Final Upper Bound =', b, '\n')
  estimated.minimizer = (a + b)/2
  cat('Estimated Minimizer =', estimated.minimizer, '\n')
}
golden.section.search(f,0,3,0.0000001)
```  
*Unconstrained optimization in multiple dimensions $(f(x1, x2) = exp(0.1 * ((x2 - x1^2))^2 + 0.05*(1 - x1)^2)$ using starting point x(0) = [-0.3, 0.8] and the default tolerance for the convergence test 10-6.)* 

**Steepest Descent Method**  
```{r}
library("pracma")
dummy <- function(x)
{
  z <- x[1]
  y <- x[2]
  rez <- exp(0.1 * ((y - z^2))^2 + 0.05*(1 - z)^2)
  rez
}
n <- 0          
eps <- 1        
a <- 0.09       
x <- c(-0.3,0.8)  
#Computation loop 
while (eps > 1e-10  && n<100)
{
gradf <- grad(dummy,x)   
eps <- abs(gradf)+abs(gradf)                                              
y <- x- a * gradf                                                                      
x <- y                                                                              
n <- n+1                                                                            
} 
#display end values
print(n)
print(x)
print(eps)

```

**Newton Method for Unconstrained optimisation in n dimension**
```{r}
library("pracma")
dummy <- function(x)
{
  z <- x[1]
  y <- x[2]
  rez <- exp(0.1 * ((y - z^2))^2 + 0.05*(1 - z)^2)
  rez
}
n <- 0                  
eps <- 1                
x <- c(1,1)           

#Computation loop 
while (eps>1e-10 && n<100 )
{
  gradf <- grad(dummy , x )                                                      
  eps <- abs(gradf)+abs(gradf)                                                   
  Hf <- hessian(dummy,x)
  k <- solve(Hf,-gradf) 
  y <- x + k
  x <- y                                                                         
  n <- n+1                                                                        
}
print(n)
print(x)

```
**Quasi-Newton for unconstrained optimisation in n dimension**
```{r}
library("pracma")
dummy <- function(x)
{
  z <- x[1]
  y <- x[2]
  rez <- exp(0.1 * ((y - z^2))^2 + 0.05*(1 - z)^2)
  rez
}
x <- c(-0.3,0.8) 
a <- 0.09
B0 <- hessian(dummy, x)
while (eps>1e-10 && n<100 )
{
  grad1 <- grad(dummy,x)
  eps <- abs(grad1)+abs(grad1)                                                   
  Bk <- hessian(dummy,y)
  p <- solve(Bk, -grad1)
  s <- a * p
  y <-  x + s
  x <- y
  grad2 <- grad(dummy, y)
  yk <- grad2 - grad1
  ykt <- transpose(yk)
  Bk1 <- (Bk + (yk * ykt)/(ykt * s) - (Bk * s)* transpose((Bk * s))/(transpose(s) * Bk* s))
  n <- n+1
}
print(n)
print(y)
```

**Direct search methods (Nelder-Mead simplex direct search) $(f(x1, x2) = ((x1^2+x2-11)^2 +((x1+x2^2-7)^2)$ using starting point x(0) = [0,-2].**
```{r}
 library("neldermead")
banana <- function(x){
  z <- x[1]
  y <- x[2]
  rez <- exp(0.1 * ((y - z^2))^2 + 0.05*(1 - z)^2)
  rez
}
opt <- optimset(MaxIter=10)
sol <- fminsearch(banana, c(0,-2), opt)
# Final Answer
sol
```


