---
title: "Snake_and_Ladders"
author: "Saurav Singh"
date: "19 February 2017"
output: pdf_document
---

## Snake & Ladders Solution
```{r}
# Snakes and ladders Simulation
```{r}
library("ggplot2")
ladder.df <- data.frame(start=c(3,11), end=c(13,17))
slide.df <- data.frame(start=c(10,16,18), end=c(5,2,8))
#out <- for (i in c(1:num.iter)) {
out <-function()
{
curLoc <- 0 # Current location
nroll <- 0 # Number of rolls
slides <- 0 # Number of slides encountered
ladders <- 0 # Number of ladders encountered

# Keep rolling dice and moving until reach 100 or greater ending the game
while(curLoc < 20) {
  roll <- sample(6,size = 1,replace = TRUE,prob = c(1/6,1/6,1/6,1/6,1/6,1/6)) # generate random number between [1 to 6]
  curLoc <- curLoc + roll # increase position
  nroll <- nroll + 1 # increase number of rolls
  # Need to check if we landed on a ladder or slide and move forward or back
  if (any(ladder.df$s %in% curLoc)) {
    curLoc <- ladder.df$e[ladder.df$s %in% curLoc]
    ladders <- ladders + 1
  }
  if (any(slide.df$s %in% curLoc)) {
    curLoc <- slide.df$e[slide.df$s %in% curLoc]
    slides <- slides + 1
  }
  out.info <- list(No_of_rolls=nroll, No_of_ladder=ladders, No_of_snakes=slides)
  
  
}
return (nroll)
}
# Expectation of N
plays<-replicate(1000,out())
mean(plays)
freq<-table(plays)
freq
percent<-freq/10
percent
frq = as.data.frame(percent)
frq
plot(frq)
```