---
title: "Assignment -1"
author: "Saurav Singh"
date: "29 January 2017"
output: pdf_document
---

#Introduction to Linear and Non Linear Optimization.
##A.Finding the roots of f(x)=0.  
1. Naive Approach ($f(x)=2^{x}-x^{2}$ for x belongs to [-2,5])  
```{r}
par(mfrow=c(2,2))
curve(2^x-x^2,-2,5)
```

* graphical solution  
```{r}
par(mfrow=c(2,2))
curve(2^x-x^2,-0.75,0)
```

So,The first root is -0.75
```{r}
par(mfrow=c(2,2))
curve(2^x-x^2,1,1.97)
```

The second root is 1.97
```{r}
par(mfrow=c(2,2))
curve(2^x-x^2,4,4.2)
```

And the last root is 4.00

* random search
```{r}
f<-function(x){2^x-x^2}
randomsearch<-function(a,b,f){
  for(i in range(a,b)){
    c<-(a+b)/2
    if (f(c)*f(a)<0){
      b<-c
    }
    else
    {
      a<-c
    }
  }
  c
}
randomsearch(-2,0,f)
randomsearch(0,3,f)
randomsearch(3,5,f)
```
  
2. Bracketing ($f(x)=cos(1/x^{2})$ for x belongs to [0.3,0.9])
```{r}
par(mfrow=c(2,2))
curve(cos(1/x^2),0.3,0.9)
f<- function(x){cos(1/x^2)}
bracket<-function(a,b,f,iter=400){
  for(i in c(1:iter)){
    c<-(a+b)/2
    if (f(c)*f(a)<0){
      b<-c
    }
    else
    {
      a<-c
    }
  }
  c
}
bracket(0.3,0.34,f)
bracket(0.34,0.4,f)
bracket(0.4,0.6,f)
bracket(0.6,0.9,f)
```
  
3. Bisection ($f(x)=cos(1/x^{2})$ for x belongs to [0.3,0.9])  
As We can see from the graph it has 4 roots.
```{r}
bisection<-function (f, a, b, tol = 0.001, m = 100) 
{
    iter <- 0
    f.a <- f(a)
    f.b <- f(b)
    while (abs(b - a) > tol) {
        iter <- iter + 1
        if (iter > m) {
            warning("maximum number of iterations exceeded")
            break
        }
        xmid <- (a + b)/2
        ymid <- f(xmid)
        if (f.a * ymid > 0) {
            a <- xmid
            f.a <- ymid
        }
        else {
            b <- xmid
            f.b <- ymid
        }
    }
    root <- (a + b)/2
    return(root)
}
par(mfrow=c(2,2))
curve(cos(1/x^2),0.3,0.9)
f <- function(x) {cos(1/x^2)}
bisection(f,0.3,0.33)
bisection(f,0.33,0.4)
bisection(f,0.4,0.6)
bisection(f,0.6,0.9)
```
  
4. Fixed Point Method ($f(x)=x-x^{7/5}+1/5$ for x~0~=0.2)
```{r}
fixedpoint <- function(fun, x0, tol=1e-07, niter=500){
  xold <- x0
  xnew <- fun(xold)
  for (i in 1:niter) {
    xold <- xnew
    xnew <- fun(xold)
    if ( abs((xnew-xold)) < tol )
      return(xnew)
  }
  stop("exceeded allowed number of iterations")
}
par(mfrow=c(2,2))
curve(x-x^(7/5)+1/5,0,2)
f <- function(x) {x - x^7/5 + 1/5}
gfun <- function(x) {x-x+x^7/5 + 1/5}
x=fixedpoint(gfun,0.2)
f(x)
```
  
5. Newton's Method ($f(x)=e^{-x}log(x)-x^{2}+x^{3}/3+1$ for x~0~=2.750,0.805,0.863 and 1.915)
```{r}
newton<-function (f, fp, x, tol = 0.001, m = 100) 
{
    iter <- 0
    oldx <- x
    x <- oldx + 10 * tol
    while (abs(x - oldx) > tol) {
        iter <- iter + 1
        if (iter > m) 
            stop("No solution found")
        oldx <- x
        x <- x - f(x)/fp(x)
    }
    return(x)
}
par(mfrow=c(2,2))
curve(exp(-x)*log(x)-x^2+((x^3)/3)+1,0.1,3)
f <- function(x) {exp(-x)*log(x)-x^2+((x^3)/3)+1}
fp <- function(x) {-exp(-x)*log(x)-2*x+x^2+exp(-x)/x}
newton(f,fp,2.750)
newton(f,fp,0.805)
newton(f,fp,0.863)
```
  
##B.Solving system of linear Equations.
Consider the matrix A, the vector b, and the solution of the linear system Ax = b given below.
```{r}
A<-matrix(floor(rnorm(25,1,2)),5,5)
b<-c(floor(rnorm(5)))
A
b
solve(A,b)
```
  
Direct Methods

1. LU Factorization.
```{r}
A <- matrix( c ( 1, 2, 2, 1 ), nrow=2, byrow=TRUE)
B <- matrix( c( 2, 6, 3, 8 ), nrow=2, byrow=TRUE )

luDecomposition <-function (A) 
  {
    nCOL <- ncol(A)
    U <- matrix(rep(0, each = nCOL * nCOL), nrow = nCOL, byrow = T)
    L <- matrix(rep(0, each = nCOL * nCOL), nrow = nCOL, byrow = T)
    U[upper.tri(A, diag = TRUE)] <- A[upper.tri(A, diag = TRUE)]
    L[lower.tri(A, diag = FALSE)] <- A[lower.tri(A, diag = FALSE)]
    diag(L) <- 1
    return(list(U = U, L = L))
}
luA <- luDecomposition(A)
L <- luA$L
U <- luA$U
print( L )
print( U )
y <- solve(L,B)
print( y )
x <- solve(U,y)
print( x )
```

2. Cholesky Factorization.
``` {r}
A<-matrix(floor(rnorm(25,1,2)),5,5)
B<-c(floor(rnorm(5)))
print(A)

x <- t(A)%*%A    
y <- t(A)%*%B   
cholesky <- function (A, tol = 1e-07) 
{
  nROW <- ncol(A)
  L <- matrix(rep(0, each = nROW * nROW), nrow = nROW, byrow = T)
  for (i in 1:nROW) {
    Aii <- A[i, i] - sum(L[i, 1:i] * L[i, 1:i])
    if (Aii < 0) {
      stop("Matrix no positive definate")
    }
    else {
      L[i, i] <- sqrt(Aii)
    }
    if ((i + 1) <= nROW) {
      for (k in (i + 1):nROW) {
        L[k, i] <- (A[k, i] - sum(L[k, 1:i] * L[i, 1:i]))/L[i, i]
      }
    }
  }
  return(L)
}
c <- cholesky(x)
print(x)
print(y)
k <- print(t(c)) 
z <- solve(k,y) 
print(z)          
l <- solve(c,z) 
print(l)         
```

3. QR Decomposition.
```{r}
hilbert <- function(n) { i <- 1:n; 1 / outer(i - 1, i, "+") }
h9 <- hilbert(9); h9
qr(h9)$rank         
qrh9 <- qr(h9, tol = 1e-10)
qrh9$rank             
y <- 1:9/10
x <- qr.solve(h9, y, tol = 1e-10) 
x <- qr.coef(qrh9, y) 
h9 %*% x            
```

4. Singular Value Decomposition.
``` {r}
A<-matrix(floor(rnorm(25,1,2)),5,5)
B<-c(floor(rnorm(5)))
k <- function (x, nu = min(n, p), nv = min(n, p), LINPACK = FALSE) 
{
    x <- as.matrix(x)
    if (any(!is.finite(x))) 
        stop("infinite or missing values in 'x'")
    dx <- dim(x)
    n <- dx[1L]
    p <- dx[2L]
    if (!n || !p) 
        stop("a dimension is zero")
    La.res <- La.svd(x, nu, nv)
    res <- list(d = La.res$d)
    if (nu) 
        res$u <- La.res$u
    if (nv) {
        if (is.complex(x)) 
            res$v <- Conj(t(La.res$vt))
        else res$v <- t(La.res$vt)
    }
    res
}

asvd <- k(A)
print(asvd)
adiag <- diag(1/asvd$d)
print(adiag)
adiag[3,3] = 0
solution = asvd$v %*% adiag %*% t(asvd$u) %*% B
print(solution)
check <- A %*% solution
# final Answer
print(check)
```


Iterative Methods.


1. Jacobi
```{r}
a <- matrix(c(2,1,5,7), nrow=2, byrow = TRUE)
b <- matrix(c(11, 13), nrow=2, byrow = TRUE)
jacobi <- function (a, b, e = 0.001) 
{
T<- array(0,dim=c(5,1))
 n<-5
 l<-0
  for (i in c(1:n))
    T[i][0]<-0
  while (l!=n)
  {
    l<-0
    for (i in c(1:n))
    {
      x[i][0]<-(1/a[i][i])*(b[i][0]);
      for (j in c(1:n))
      {
        if (j!=i)
          x[i][0] <- x[i][0]-(1/a[i][i])*(a[i][j]*T[j][0]);
      }
    }
    for(i in c(1:n))
    {
      k<-abs(x[i][0]-T[i][0]);
      if (k<=e)
      {
        l<-l+1;
      }
    }
    for (i in c(1:n))
      T[i][0] <- x[i][0];
  }
  for (i in c(1:n))
    print(x[i][0])
}
```
**Output**  
x1=7.11096  
x2=-3.22174  

2. Gauss-Seidel
```{r}
a<-matrix(floor(rnorm(25,1,2)),5,5)
b<-c(floor(rnorm(5)))
x <-c(0,0,0,0,0)
Seidel <- function(a , b, x)
{
n <- 5
m <- 5
i <- 0
j <- 0
y<- array(0,dim=c(0,5))

  while (m > 0)
  {
    for (i in c(1:n))
    {
      y[i] <- (b[i] / a[i][i])
      for (j in c(1:n))
      {
        if (j != i)
        y[i] <- y[i] - ((a[i][j] / a[i][i]) * x[j]);
        x[i] <- y[i];
      }
       
    }
    m <- m - 1
  }
return (y)
}

```
**Output**  
x1=-1.33333  x2=0.33333  x3=-7.33333   x4=-1.4444  x5=nan  
x1=nan  x2=nan  x3=nan  x4=nan  x5=nan  
x1=nan  x2=nan  x3=nan  x4=nan  x5=nan  

3. Success Overelaxation Method
```{r}
library("optR")
a<-matrix(floor(rnorm(25,1,2)),5,5)
b<-c(floor(rnorm(5)))
SOR<- function(a,b,w = 1.3, tol = 1e-07)
{
  n <-5
  D <- diag(a)
  luA <- LUsplit(a)
  L <- luA$L
  U <- luA$U
  e <- max(eigen(inv.optR(D+w*L ) * ( D*(1-w) - w*U)))
  if(abs(e)>1)
  {
    print("Since the modulus of largest eigen value of iterative matrix is not less than 1")
    print("The Process is not convergent")
    return 
  }
  err <- 10000000 * runif(1,5.0,1)
  x <-c(0,0,0,0,0)
  p <- 0
  v <- matrix(c(0,0,0,0,0), nrow = 5, byrow = TRUE)
  while((sum(abs(err)) >= tol) == v) 
  {
    for (i in c(1:n))
    {
      for ( j in c(1:n))
        if( j != i)
          p <- p + a[i][j] * x[j]
    }
    x[i] <- (1 - w) * x[i] + (w/a[i][i]) * (b[i]-p)
  }
  for (i in c(1:n))
    print(x[i][0])
}
```

4. Block Iterative Method
```{r}
a <- matrix(c(2,1,5,7), nrow=2, byrow = TRUE)
b <- matrix(c(11, 13), nrow=2, byrow = TRUE)
jacobi <- function (a, b, e = 0.001) 
{
T<- array(0,dim=c(5,1))
 n<-5
 l<-0
  for (i in c(1:n))
    T[i][0]<-0
  while (l!=n)
  {
    l<-0
    for (i in c(1:n))
    {
      x[i][0]<-(1/a[i][i])*(b[i][0]);
      for (j in c(1:n))
      {
        if (j!=i)
          x[i][0] <- x[i][0]- T[j][0]
      }
    }
    for(i in c(1:n))
    {
      k<-abs(x[i][0]-T[i][0]);
      if (k<=e)
      {
        l<-l+1;
      }
    }
    for (i in c(1:n))
      T[i][0] <- x[i][0];
  }
  for (i in c(1:n))
    print(x[i][0])
}
```

