---
title: "Assignment -1"
author: "Saurav Singh"
date: "29 January 2017"
output: pdf_document
---

#Introduction to Linear and Non Linear Optimization.
##A.Finding the roots of f(x)=0.  
1. Naive Approach ($f(x)=2^{x}-x^{2}$ for x belongs to [2-,5])  

* graphical solution  
* random search
```{r}
par(mfrow=c(2,2))
curve(2^x-x^2,-2,5)
```
  
2. Bracketing ($f(x)=cos(1/x^{2})$ for x belongs to [0.3,0.9])
```{r}
par(mfrow=c(2,2))
curve(cos(1/x^2),0.3,0.9)
```
  
3. Bisection ($f(x)=cos(1/x^{2})$ for x belongs to [0.3,0.9])  
As We can see from the graph it has 4 roots.
```{r}
library(cmna)
bisection
par(mfrow=c(2,2))
curve(cos(1/x^2),0.3,0.9)
f <- function(x) {cos(1/x^2)}
bisection(f,0.3,0.33)
bisection(f,0.33,0.4)
bisection(f,0.4,0.6)
bisection(f,0.6,0.9)
```
  
4. Fixed Point Method ($f(x)=x-x^{7/5}+1/5$ for x~0~=0.2)
```{r}
par(mfrow=c(2,2))
curve(x-x^(7/5)+1/5,0,2)
```
  
5. Newton's Method ($f(x)=e^{-x}log(x)-x^{2}+x^{3}/3+1$ for x~0~=2.750,0.805,0.863 and 1.915)
```{r}
newton
par(mfrow=c(2,2))
curve(exp(-x)*log(x)-x^2+((x^3)/3)+1,0.1,3)
f <- function(x) {exp(-x)*log(x)-x^2+((x^3)/3)+1}
fp <- function(x) {-exp(-x)*log(x)-2*x+x^2+exp(-x)/x}
newton(f,fp,2.750)
newton(f,fp,0.805)
newton(f,fp,0.863)
```
  
##B.Solving system of linear Equations.
Consider the matrix A, the vector b, and the solution of the linear system Ax = b given below.
```{r}
A<-matrix(floor(rnorm(25,1,2)),5,5)
b<-c(floor(rnorm(5)))
A
b
solve(A,b)
```
  
Direct Methods

1. LU Factorization.
```{r}
A <- matrix(c(1, 2, -7, -1, -1, 1, 2, 1, 5), 3)
lumatrix(A)
```
2. Cholesky Factorization.
```{r}
(A <- matrix(c(5, 1, 2, 1, 9, 3, 2, 3, 7), 3))
A
(L <- choleskymatrix(A))
t(L) %*% L
```
3. QR Decomposition.
4. Singular Value Decomposition.


Iterative Methods.

1. Jacobi
```{r}
A<-matrix(floor(rnorm(25,1,2)),5,5)
b<-c(floor(rnorm(5)))
A
b
jacobi(A, b)
```
2. Gauss-Seidel
```{r}
gaussseidel(A,b)
```
3. Success Overelaxation Method

4. Block Iterative Method